---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Question 5: Volatility Modelling"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Ruan Geldenhuys"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch University, Stellenbosch, South Africa" # First Author's Affiliation
Email1: "22550801\\@sun.ac.za" # First Author's Email address

# Author2: "John Smith"
# Ref2: "Some other Institution, Cape Town, South Africa"
# Email2: "John\\@gmail.com"
# CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.
# 
# Author3: "John Doe"
# Email3: "Joe\\@gmail.com"
# 
# CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
# keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
# JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
# BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
# addtoprule: TRUE
# addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
#Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
#abstract: |
  
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf.
# These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!

# Lets load in example data, and see how this can be stored and later called from your 'data' folder.
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
library(fmxdat)
pacman::p_load("MTS", "robustbase")
pacman::p_load("tidyverse", "devtools", "rugarch", "rmgarch", 
    "forecast", "tbl2xts", "lubridate", "PerformanceAnalytics", 
    "ggthemes")

```

# Introduction

This question investigates the Rand's notorious volatility. To this I first fit a univariate GARCH in order to investigate the structural volatility of the Rand. I then compare it with other highly volatile currencies. Finally, I fit a multivariate GARCH between the Rand and a G10 currency basket. It appears that, although the Rand is volatile, it is far from the most volatile currency.

```{r}
cncy <- read_rds("data/currencies.rds")
cncy_Carry <- read_rds("data/cncy_Carry.rds")
cncy_value <- read_rds("data/cncy_value.rds")
cncyIV <- read_rds("data/cncyIV.rds")
bbdxy <- read_rds("data/bbdxy.rds")

```


```{r}

cncy_ret <- cncy %>% 
    group_by(Name) %>% 
    mutate(Returns = Price/lag(Price) - 1) %>% 
    filter(date > first(date)) %>% 
    mutate(Name = gsub("\\_Cncy", "", Name)) %>% 
    mutate(Name = gsub("\\_Inv", "", Name)) 

sa_cncy <- cncy_ret %>% 
    filter(Name == 'SouthAfrica') %>% 
    select(c(date, Returns))
sa_cncy_2 <- sa_cncy[,2:3]   
sa_xts_returns <- tbl_xts(sa_cncy_2)
```

```{r}
garch1 <- ugarchspec(
    variance.model = list(model = c("sGARCH","gjrGARCH","eGARCH","fGARCH","apARCH")[1], 
    garchOrder = c(1, 1)), 
    mean.model = list(armaOrder = c(1, 0), include.mean = TRUE), 
    distribution.model = c("norm", "snorm", "std", "sstd", "ged", "sged", "nig", "ghyp", "jsu")[1])


garchfit1 = ugarchfit(spec = garch1,data = sa_xts_returns) 
```

```{r Figure1,  warning =  FALSE, fig.align = 'center', fig.cap = "Conditional SD of Rand", fig.ext = 'png', fig.height = 3, fig.width = 6}

sigma <- sigma(garchfit1) %>% xts_tbl() 
colnames(sigma) <- c("date", "sigma") 
sigma <- sigma %>% mutate(date = as.Date(date))

plot(garchfit1, which = 3)

```

The plot above shows the structural volatility of the Rand compared to the noise.

```{r}
sd_df <- cncy_ret %>% 
    group_by(Name) %>% 
    summarise(SD = sd(Returns)) %>% 
    arrange(desc(SD)) %>% 
    head(10)

top10 <- unique(sd_df$Name)

library(kableExtra)
VolatilityTable <- kable(sd_df)
VolatilityTable
```

The table above shows the top 10 most volatile currencies ranked. South Africa is ranked 10th. While not the most volatile, it is certainly high, considering we have 41 currencies in the dataset. Nigeria takes the top spot with a SD of 0.0383.


```{r Figure2,  warning =  FALSE, fig.align = 'center', fig.cap = "Conditional Standard Deviations Over Time", fig.ext = 'png', fig.height = 3, fig.width = 6}
top10_returns <- cncy_ret %>% 
    filter(Name %in% top10) %>% 
    select(c(date, Name, Returns))


garches <- top10_returns %>%
  group_by(Name) %>%
  group_split() %>%  # Split the data into a list of data frames, one for each group
  map(~ {
    data_xts = xts(.x$Returns, order.by = as.Date(.x$date))
    ugarchfit(spec = garch1, data = data_xts)
  })


sigma_list <- map(garches, ~ data.frame(Date = index(sigma(.x)), Sigma = sigma(.x)))

# Optional: Add a column for the 'Name' identifier
names_list <- top10_returns$Name %>% unique()
sigma_df <- map2_df(sigma_list, names_list, ~ mutate(.x, Name = .y))

# View the combined data frame
g1 <- ggplot(sigma_df, aes(x = Date, y = Sigma, color = Name)) +
  geom_line() +
  ylim(c(0,0.2))+
  labs(title = "Conditional Standard Deviations Over Time",
       x = "Date",
       y = "Sigma",
       color = "Group") +
  theme_fmx() +
  theme(legend.position = "bottom")
finplot(g1)
```

The plot above shows the conditional standard deviations over time for the top 10 most volatile countries. What is immediately apparent is that every currency experience volatility spikes. However, these spikes occur at different times for different countries, indicating that the structural volatilities don't tend to follow each other.

```{r Figure3,  warning =  FALSE, fig.align = 'center', fig.cap = "Dynamic Conditional Correlations: ZAR and G10", fig.ext = 'png', fig.height = 3, fig.width = 6}
DBHVG10U<- cncy_Carry %>% 
    filter(Name == "DBHVG10U") %>% 
    mutate(g10_Return = Price/lag(Price)-1) %>% 
    filter(date > first(date)) %>% 
    select(c(date, g10_Return))

joined_df <- left_join(DBHVG10U, sa_cncy_2, by = 'date') %>% 
    rename(sa_xts_returns = 'Returns')  %>% 
    tbl_xts()
    
DCCPre <- dccPre(joined_df, include.mean = T, p = 0)
StdRes <- DCCPre$sresi
DCC <- dccFit(StdRes, type="Engle")

Rhot <- DCC$rho.t

dateVector <- xts_tbl(joined_df)
dateVector <- dateVector %>% 
    select(date)
plotting_df <- cbind(dateVector, Rhot[,2]) 
colnames(plotting_df)<- c('date', 'Rhot')

g2 <- ggplot(plotting_df, aes(x=date, y=Rhot))+
    geom_line(color = "blue")+
    labs(title = 'Dynamic Conditional Correlations: ZAR & G10',
         y= 'rho',
         x= 'Date')+
    theme_hc()
g2

```
The plot above shows the dynamic conditional correlation of the ZAR and the G10 basket. 

```{r}
#Attempted this - cant get it to work

# top4_and_SA <- c("Nigeria", "Romania", "Bulgaria", "Zambia", "SouthAfrica")
# multivariateDF <- cncy_ret %>% 
#     filter(Name %in% top4_and_SA) %>%
#     select(c(date, Name, Returns)) %>% 
#     pivot_wider(names_from = Name,
#                 values_from = Returns) %>%
#     filter(complete.cases(.)) %>% 
#     tbl_xts()
# 
# DCCPre <- dccPre(multivariateDF, include.mean = T, p = 0)


```



