---
output:
  md_document:
    variant: markdown_github
---

# Purpose

Purpose of this work folder.

Ideally store a minimum working example data set in data folder.

Add binary files in bin, and closed R functions in code. Human Readable settings files (e.g. csv) should be placed in settings/


```{r include=FALSE}

rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.
library(tidyverse)
library(fmxdat)
library(RcppRoll)
library(fmxdat)
library(RColorBrewer)
library(tbl2xts)
library(PerformanceAnalytics)
library(ggExtra)
library(kableExtra)
```


## Question 1

This question investigates returns from an AI Fund, a market benchmark (CAPPED SWIX) and active managers. It aims to shed light on the differences and similarity in return structures through various visualisation techniques.

In terms of data preparation, I calculate the average returns for both ASISA and the benchmark across all funds and tickers. I then join these returns with the returns of the AI fund into one data frame. I then convert this into an 'xts' object which allows me to use the 'PerformanceAnalytics' package.

```{r }
#Load and prep the data
ASISA <- read_rds("data/ASISA_Rets.rds")
BM <- read_rds("data/Capped_SWIX.rds")
AI_Fund <- read_rds("data/AI_Max_Fund.rds")

#Calculate the average ASISA return across all funds for a given month
ASISA_average <- ASISA %>% 
    group_by(date) %>% 
    summarise(meanReturn = mean(Returns))

#Calculate the average BM return across all indices for a given month
BM_average <- BM %>% 
    group_by(date) %>% 
    summarise(meanReturn = mean(Returns))


#Get all the returns in the same dataframe
returnDF <- left_join(ASISA_average, BM_average, by = 'date') %>% 
    left_join(AI_Fund, by='date')
colnames(returnDF)<- c('date','Active', 'Benchmark', 'AI')

returnDF <- returnDF %>% 
    pivot_longer(cols = c(-date),
                 names_to = "Fund",
                 values_to = "Returns")

#Convert to XTS
xts.returnDF <- returnDF %>% 
    tbl_xts(tblData = ., cols_to_xts = Returns, spread_by = Fund)

```

```{r}
AIvsActive <- chart.Scatter(x = xts.returnDF$AI, 
                            y = xts.returnDF$Active, 
                            main = "Scatter: AI vs Active Managers", col = "darkred",symbolset = 16, 
                            xlab = "AI", ylab = "Active Managers")
```

The figure above, made by the 'PerformanceAnalytics' package, shows a scatter between the returns of active managers and the AI fund. A clear positive correlation is visible. 

```{r}
AIvsBenchmark <- chart.Scatter(x = xts.returnDF$AI, 
                            y = xts.returnDF$Benchmark, 
                            main = "Scatter: AI vs Benchmark", col = "darkblue",symbolset = 16, 
                            xlab = "Benchmark", ylab = "Benchmark Indices")
```

The figure above now shows a scatter between the returns of the benchmark and the AI fund. An even stronger positive correlation is now visible. This means that for, in both cases, it is rare for the AI fund to show a negative return when the other funds show a positive return and vice-versa.

The next step is to calculate rolling returns for all 3 funds. I calculate the annualised rolling 3 year return using the 'roll_prod' function in the 'RcppRoll' package. I then join these rolling returns together and then convert this data frame into long format to get it ready for plotting.

```{r}
##ASISA
#Calculate the annualized three year rolling return
ASISA_RollRet <- ASISA_average %>% 
    mutate(RollRets_ASISA = RcppRoll::roll_prod(1 + meanReturn, 36, fill = NA, align = "right")^(12/36)-1) %>% 
    group_by(date) %>% 
    filter(any(!is.na(RollRets_ASISA))) %>% 
    ungroup() %>%
    select(date, RollRets_ASISA)
    
##BM
##Calculate the annualized three year rolling return
BM_RollRet <- BM_average %>% 
    mutate(RollRets_BM = RcppRoll::roll_prod(1 + meanReturn, 36, fill = NA, align = "right")^(12/36)-1) %>% 
    group_by(date) %>% 
    filter(any(!is.na(RollRets_BM))) %>% 
    ungroup() %>%
    select(date, RollRets_BM)

joinedDF <- left_join(ASISA_RollRet, BM_RollRet, by = "date")

##AIFund
#Calculate the annualized three year rolling return
AI_RollRet <- AI_Fund %>% 
    mutate(RollRets_AI = RcppRoll::roll_prod(1 + AI_Fund, 36, fill = NA, align = "right")^(12/36)-1) %>% 
    group_by(date) %>% 
    filter(any(!is.na(RollRets_AI))) %>% 
    ungroup() %>%
    select(date, RollRets_AI)

joinedDF <- joinedDF %>% 
    left_join(AI_RollRet, by = "date")
colnames(joinedDF) <- c('date','Active','Benchmark','AIFund')


#Creating plotting dataframe
plottingDF <- joinedDF %>% 
    pivot_longer(cols = c(-date),
                 names_to = "Fund",
                 values_to = "RollingReturns")

```


```{r warning=FALSE}
g <- plottingDF %>% 
    ggplot() + 
    geom_line(aes(date, RollingReturns, color = Fund), alpha = 0.7, size = 1.25) + 
    labs(title = "Rolling 3 Year Annualized Returns Comparison", 
        subtitle = "", x = "", y = "Rolling 3 year Returns (Ann.)", 
        caption = "") + 
    theme_fmx(title.size = ggpts(30), subtitle.size = ggpts(5), caption.size = ggpts(25), CustomCaption = F) + 
    fmx_cols()+
    scale_fill_brewer(palette = "Accent") 
    

finplot(g, x.date.dist = "1 year", x.date.type = "%Y", x.vert = T, 
    y.pct = T, y.pct_acc = 1)
```

The rolling 3 year annualised returns of all three funds are shown above. All three funds tend to follow each other and show significant downturns during the GFC and Covid-19. Comparatively, the active managers perform the worst, as for large periods of time there rolling returns are substantially lower than the other two funds. The AI fund seems capable of outperforming the benchmark, but never for long periods of time.

Next, I pivot the returns dataframe wider in order to apply a 25bps fee on both the active managers and the AI fund. I then take the dataframe back to long format in order to once again get it ready for plotting.

```{r}
returnDF <- returnDF %>% 
    pivot_wider(names_from = Fund, values_from = Returns)
returnDF <- returnDF %>% 
    mutate(Active = Active - 0.025/12) %>%   # Applying a 25bps fee for active and AI fund
    mutate(AI = AI - 0.025/12)
returnDF <- returnDF %>% 
    pivot_longer(cols = c(-date),
                 names_to = "Fund",
                 values_to = "Returns")
```

```{r warning=FALSE}
# Plot the density functions of each return and add a line for the mean benchmark return
mean_benchmark_returns <- mean(returnDF$Returns[returnDF$Fund == "Benchmark"])

densityPlot <- ggplot(returnDF, aes(x = Returns)) + 
    geom_density(aes(fill = Fund), alpha = 0.5, color = NA) +
    geom_vline(xintercept = mean_benchmark_returns, linetype = "dashed", color = "black") +
    geom_text(aes(x = mean_benchmark_returns, y = 15), label = "Market Benchmark", 
              vjust = -1, hjust = -0.05, color = "black") +
    labs(title = "Density of Returns by Fund",
         x = "Returns",
         y = "Density") +
    theme_minimal() +
    scale_fill_brewer(palette = "Accent") +
    theme(panel.border = element_rect(colour = "black", fill=NA, size=1))
densityPlot
```
The density functions of returns, grouped by fund, show a clear result. Active managers tend to perform below the market benchmark. The marker benchmark here was defined as the average return of the benchmark, as can be seen in the code above. The market benchmark and the AI fund perform very similar. The key difference here lies in the flatter distribution of the AI fund, meaning it is less consistent that the market benchmark.

The next chunk of code aims to compare the AI fund with the best and worst performing actively managed fund. This is done by first filtering the raw data to contain only funds with complete data, as some funds only started recently or closed down, which makes comparison difficult. Next the best active fund is selected by calculating the average return for all funds and selecting the best one. The worst fund is found the exact same way. I then calculate the rolling returns for these two funds, using 'RcppRoll' and then join them in a data frame with the rolling returns for the AI fund

```{r}

#Filter out funds that don't have complete date
years <- 2003:2023
funds_with_complete_data <- ASISA %>%
  mutate(Year = year(as.Date(date))) %>%   
  group_by(Fund) %>%
  summarize(All_Years_Present = all(years %in% unique(Year))) %>% 
  filter(All_Years_Present) %>%
  pull(Fund)

#Find best performing fund
BestActive <- ASISA %>%
  filter(Fund %in% funds_with_complete_data) %>%
  group_by(Fund) %>%
  summarise(Average_Return = mean(Returns, na.rm = TRUE)) %>%
  arrange(desc(Average_Return)) %>%
  slice(1)

BestActiveReturns <- ASISA %>% 
    filter(Fund == BestActive$Fund) %>% 
    select(c('date','Returns'))


##Find worse performing fund
WorstActive <- ASISA %>%
  filter(Fund %in% funds_with_complete_data) %>%
  group_by(Fund) %>%
  summarise(Average_Return = mean(Returns, na.rm = TRUE)) %>%
  arrange(Average_Return) %>%
  slice(1)

WorstActiveReturns <- ASISA %>% 
    filter(Fund == WorstActive$Fund) %>% 
    select(c('date','Returns'))

#Get the rolling returns for best and worst fund
Best_RollRet <- BestActiveReturns %>% 
    mutate(RollRets_best = RcppRoll::roll_prod(1 + Returns, 36, fill = NA, align = "right")^(12/36)-1) %>% 
    group_by(date) %>% 
    filter(any(!is.na(RollRets_best))) %>% 
    ungroup() %>%
    select(date, RollRets_best)

Worst_RollRet <- WorstActiveReturns %>% 
    mutate(RollRets_worst = RcppRoll::roll_prod(1 + Returns, 36, fill = NA, align = "right")^(12/36)-1) %>% 
    group_by(date) %>% 
    filter(any(!is.na(RollRets_worst))) %>% 
    ungroup() %>%
    select(date, RollRets_worst)

##Group these funds and AI into one dataset
BestAndWorseDF <- left_join(Best_RollRet, Worst_RollRet, by = 'date') %>% 
    left_join(AI_RollRet, by = 'date')
colnames(BestAndWorseDF) <- c('date', 'BestActive', 'WorstActive', 'AI')
```
Next I plot these funds and shade the area between the best and worse fund.

```{r}
plotDiff <- ggplot(BestAndWorseDF, aes(x = date)) +
  geom_line(aes(y = BestActive, colour = "BestActive")) +
  geom_line(aes(y = WorstActive, colour = "WorstActive")) +
  geom_line(aes(y = AI, colour = "AI")) +
  geom_ribbon(aes(ymin = WorstActive, ymax = BestActive), fill = "grey", alpha = 0.5) +
  labs(title = "Comparison of Best and Worst Active funds vs AI Fund",
       x = "Date",
       y = "Rolling 3 Year Return (Ann.)",
       colour = "Fund") +
  scale_colour_manual(values = c("BestActive" = "blue", "WorstActive" = "red", "AI" = "green")) +
  theme_minimal()+
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1))
plotDiff
```

The figure reveals an interesting story. It appears that the AI fund often lies between the rolling returns of these two funds. Interestingly, for a period in the middle of our sample, the worst performing fund, actually outperformed the best performing fund. This potentially speaks to the shortcomings of average returns as a metric of success. In this period, the AI fund also outperformed both of these fund. The AI fund performs lower than the worst fund at times, but this usually does not last long.

# Question 2

This question focuses on a recent article that discusses the misconception surrounding hedging out the Rand. By replicating the figure within that article I therefore show that hedging very rarely has beneficial effects. I then go further by showing how the volatility of a portfolio can actually be increased through hedging.

For data preparation I load in the data and then apply the weights specified in the question to the corresponding asset by multiplying it return with the corresponding weight. Summing these returns then yield the total portfolio returns. I then calculate the ZAR returns and lastly join the two return series' into one data frame.

```{r}
Indexes <- read_rds("data/Cncy_Hedge_Assets.rds")
ZAR <- read_rds("data/Monthly_zar.rds")

#Apply weights
Portfolio <- Indexes %>% 
    mutate(MSCI_ACWI = MSCI_ACWI * 0.18) %>% 
    mutate(Bbg_Agg = Bbg_Agg * 0.12) %>% 
    mutate(J433 = J433 * 0.42) %>% 
    mutate(ALBI = ALBI * 0.28)

#Calculate portfolio return.
PortfolioReturns <- Portfolio %>%
    rowwise() %>%
    mutate(sum = sum(c_across(-date), na.rm = TRUE)) %>%
    select(date, sum)

#Calculate ZAR returns
ZAR_Returns <- ZAR %>% 
    mutate(Return = value/lag(value) - 1) %>% 
    select(c('date','Return'))

#Join in one dataframe
joinedDF <- left_join(PortfolioReturns, ZAR_Returns, by = 'date')
colnames(joinedDF)<- c('date','Portfolio_Returns','USDZAR_Returns')
```

In order to replicate the graph in the article, the percentage of observations that fall within each quadrant needs to be calculated. This is handled by the code chunk below. First I remove the NA's from my data. I then iteratively count the number of observations in each quadrant and divide that by the total number of observations to get a percentage. I do the same for when a hedging fee is applied.

```{r}
#Calculate Quadrant Percentages
cleanJoined <- na.omit(joinedDF)
total_count <- nrow(cleanJoined)

bothPositive_count <- sum(cleanJoined$Portfolio_Returns > 0 & cleanJoined$USDZAR_Returns > 0, na.rm = TRUE)
percentage_positive <- round((bothPositive_count / total_count) * 100, 1)
percentage_positive <- paste0(percentage_positive, "%")

bothNegative_count <- sum(cleanJoined$Portfolio_Returns < 0 & cleanJoined$USDZAR_Returns < 0, na.rm = TRUE)
percentage_negative <- round((bothNegative_count / total_count) * 100, 1)
percentage_negative <- paste0(percentage_negative, "%")

PosAndNeg_count <- sum(cleanJoined$Portfolio_Returns > 0 & cleanJoined$USDZAR_Returns < 0, na.rm = TRUE)
percentage_PosAndNeg <- round((PosAndNeg_count / total_count) * 100, 1)
percentage_PosAndNeg <- paste0(percentage_PosAndNeg, "%")

NegAndPos_count <- sum(cleanJoined$Portfolio_Returns < 0 & cleanJoined$USDZAR_Returns > 0, na.rm = TRUE)
percentage_NegAndPos <- round((NegAndPos_count / total_count) * 100, 1)
percentage_NegAndPos <- paste0(percentage_NegAndPos, "%")

#Quadrant percentages after a hedging fee
TopLeft_hedging <- sum(cleanJoined$Portfolio_Returns > 0 & cleanJoined$USDZAR_Returns < -0.025, na.rm = TRUE)
percentage_TopLeft_hedge <- round((TopLeft_hedging / total_count) * 100, 1)
percentage_TopLeft_hedge <- paste0(percentage_TopLeft_hedge, "%")

BotLeft_hedging <- sum(cleanJoined$Portfolio_Returns < 0 & cleanJoined$USDZAR_Returns < -0.025, na.rm = TRUE)
percentage_BotLeft_hedge <- round((BotLeft_hedging / total_count) * 100, 1)
percentage_BotLeft_hedge <- paste0(percentage_BotLeft_hedge, "%")
```

The plot is constructed in the chunk below. (You might want to turn off warnings if you intend to run this bad boy yourself). The main plot is a scatter plot with a regression line fitted. I then add lines at 0 for both axis, and a line at -0.025 to show the fee of hedging. I then shade the quadrants using annotate. Labels for the percentages calculated above are added, as well as labels interpreting each quadrant. Lastly, distributions are added on the top and side of the graph using 'ggMarginal'

```{r warning=FALSE}
#Plotting this monster
p<- ggplot(joinedDF, aes(x = USDZAR_Returns, y = Portfolio_Returns)) +
    geom_point() +  # Add points for scatterplot
    geom_smooth(method = "lm", color = "blue") +  # Add linear regression line
    ylim(-0.2,0.2)+
    xlim(-0.2,0.2)+
    
    geom_vline(xintercept = 0, linetype = "solid") +  # Add vertical line at zero
    geom_hline(yintercept = 0, linetype = "solid") +  # Add horizontal line at zero
    
    geom_vline(xintercept = -0.025, linetype = "longdash")+
    
    annotate("rect", xmin = -Inf, xmax = 0, ymin = 0, ymax = Inf, fill = "green", alpha = 0.2) + # Top-left quadrant
    annotate("rect", xmin = 0, xmax = Inf, ymin = 0, ymax = Inf, fill = "red", alpha = 0.2) +   # Top-right quadrant
    annotate("rect", xmin = -Inf, xmax = 0, ymin = -Inf, ymax = 0, fill = "pink", alpha = 0.2) + # Bottom-left quadrant
    annotate("rect", xmin = 0, xmax = Inf, ymin = -Inf, ymax = 0, fill = "blue", alpha = 0.2) + # Bottom-right quadrant
    
    labs(title = "ZAR/USD Returns vs. Portfolio Returns",
       x = "USD/ZAR Returns",
       y = "Portfolio Returns") +
    theme_minimal()+
    
    geom_label(aes(x = -Inf, y = Inf, label = percentage_PosAndNeg), 
               hjust = 0, vjust = 1, size = 4, color = "black", fill = "white") +  # Top-left quadrant
    geom_label(aes(x = Inf, y = Inf, label = percentage_positive), 
               hjust = 1, vjust = 1, size = 4, color = "black", fill = "white") +  # Top-right quadrant
    geom_label(aes(x = -Inf, y = -Inf, label = percentage_negative), 
               hjust = 0, vjust = 0, size = 4, color = "black", fill = "white") +  # Bottom-left quadrant
    geom_label(aes(x = Inf, y = -Inf, label = percentage_NegAndPos), 
               hjust = 1, vjust = 0, size = 4, color = "black", fill = "white") +  # Bottom-right quadrant
    
    geom_label(aes(x = -0.025, y = Inf, label = percentage_TopLeft_hedge), 
               hjust = 1, vjust = 1, size = 4, color = "black", fill = "white") +
    geom_label(aes(x = -0.025, y = -Inf, label = percentage_BotLeft_hedge), 
               hjust = 1, vjust = 0, size = 4, color = "black", fill = "white") +
    
    geom_label(aes(x=-0.1, y=0.1, label = "Hedge works but \n amplifies volatility"),
               hjust = 0.5, vjust = 0.5, size = 4, color = "black", fill = "white") +
    geom_label(aes(x=0.1, y=0.1, label = "Hedge throws away returns"),
               hjust = 0.5, vjust = 0.5, size = 4, color = "black", fill = "white") +
    geom_label(aes(x=-0.1, y=-0.1, label = "Best case for Hedge"),
               hjust = 0.5, vjust = 0.5, size = 4, color = "black", fill = "white") +
    geom_label(aes(x=0.1, y=-0.1, label = "Hedge removes currency \n cushion"),
               hjust = 0.5, vjust = 0.5, size = 4, color = "black", fill = "white") +
    
     theme(panel.border = element_rect(colour = "black", fill=NA, size=1))



p <- ggMarginal(p, type = "density", fill = 'darkgreen', color = NA)

p
```

The graph interpretation can be done as follows. If we find ourselves in the top left quadrant, it means portfolio returns are positive and USD/ZAR returns are negative. In this case a hedge will work but it increases volatility, a less than desirable outcome. This occurs for 46.2% of observations and 30.1% if we apply a fee. The bottom left quadrant shows when portfolio and USD/ZAR returns are negative. This is the best case scenario for a hedge, yet it only occurs 7.5% of the time. 2.2% when applying a fee. The top right and bottom right quadrant shows outright negative positions to be in when hedging against the Rand. These occur 19.9% and 25.8% respectively.


In the code below, I first rename the columns of 'PortfolioReturns' to 'date' and 'Return', and define a set of weights for a portfolio. I then calculate the rolling standard deviation (SD) for a hedged portfolio, using returns calculated from these weights, rebalanced quarterly. This hedged portfolio's rolling SD is annualized over a 36-month window. For the unhedged portfolio, I perform a similar process but adjust the returns for currency effects using 'ZAR_Returns'. Finally, I merge the rolling SD data of both portfolios and reshape it into a long format, allowing for a comparative analysis of the rolling SDs across the two different portfolio types over time.
```{r}
colnames(PortfolioReturns) <- c('date', 'Return')

weights <- c(0.18, 0.12, 0.42, 0.28)

#Calculate Rolling SD of a hedged portfolio
port1 <- Return.portfolio(Indexes, weights = weights, rebalance_on = "quarters")

port_hedged_df <- xts_tbl(port1) %>% 
    mutate(RollSD = RcppRoll::roll_sd(1 + portfolio.returns, 36, fill = NA, align = "right") * 
    sqrt(12)) %>% 
    filter(!is.na(RollSD))

#Calculate rolling SD of an unhedged portfolio
port_unhedged_df <- xts_tbl(port1) %>% 
    left_join(ZAR_Returns, by = 'date')
colnames(port_unhedged_df) <- c('date', 'portfolioReturns', 'currencyReturns')

port_unhedged_df <- port_unhedged_df %>% 
    mutate(Unhedged_Returns = (0.95*portfolioReturns) + (0.5*currencyReturns)) %>% #Unsure if this weighting is appropriate
    select(c(date, Unhedged_Returns)) %>% 
    na.omit()

port_unhedged_df <- port_unhedged_df %>% 
    mutate(RollSD_unhedged = RcppRoll::roll_sd(1 + Unhedged_Returns, 36, fill = NA, align = "right") * 
    sqrt(12)) %>% 
    filter(!is.na(RollSD_unhedged)) 


#Join the two portfolio SD's and the convert to long
joined_portfolios <-
    left_join(port_hedged_df, port_unhedged_df, by = 'date') %>% 
    select(c(date, RollSD, RollSD_unhedged)) %>% 
    filter(!is.na(RollSD_unhedged)) %>% 
    pivot_longer(-date,
                 names_to = 'portfolio',
                 values_to = 'RollSD')
```

Below I construct a plot of the two rolling SD's of the two different portfolios.
```{r}
g <- joined_portfolios %>% 
    ggplot() + 
    geom_line(aes(date, RollSD, color = portfolio), alpha = 0.7, size = 1.25) + 
        
    labs(title = "Rolling 3 Year Annualized SD", 
        subtitle = "", x = "", y = "Rolling 3 year SD (Ann.)", 
        caption = "") + theme_fmx(title.size = ggpts(30), 
        subtitle.size = ggpts(5), caption.size = ggpts(25), CustomCaption = F) + 
        
    fmx_cols()
    
finplot(g, x.date.dist = "1 year", x.date.type = "%Y", x.vert = T, 
    y.pct = T, y.pct_acc = 1)
```

The graph displayed above illustrates the rolling three-year annualized standard deviation for both a hedged and an unhedged portfolio. What stands out is that, for most of the observed period, the unhedged portfolio exhibits lower volatility. This observation reinforces the points made earlier, demonstrating that hedging against the Rand actually elevates the portfolio's overall volatility. This outcome contradicts the primary goal of hedging against the Rand, which is to reduce volatility.

# Question 3

This questions looks at two different all share indices, namely the ALSI and the SWIX. I compare these two indices by first estimating the cumulative returns for both broken down by sector and index. Next I perform stratification analysis using the ZAR to see how these indices react during currency volatility. Lastly, I compare the differences in returns for different capped portfolios.

All functions for this question can be found in 'Questions/Question3/code/'.

```{r include=FALSE}
source('Questions/Question3/code/perf_comparisons.R')
source('Questions/Question3/code/calculatePortfolioReturns.R')
source('Questions/Question3/code/PortfolioCapping.R')

```

The code below first loads the required data. It then calculates the cumulative return for the ALSI and SWIX, first grouped by Sector and then by Index_Name. It does this by first pulling each unique sector out of the raw data. For each sector it then applies a function called 'calc_portfolio_returns'. 

This function calculates the returns for portfolios based on specified sectors and indices within a given dataset. Initially, it filters and normalizes the data based on the chosen sector and index, if provided. The function then separately processes data for two indices, J203 and J403, by selecting relevant columns, spreading the data, and converting it into a time-series format, handling any missing values. It calculates the portfolio returns for both J203 and J403 using the Safe_Return.portfolio function. Subsequently, it compiles raw data with portfolio data, calculating and summarizing portfolio returns by date. The final output joins the calculated returns from both J203 and J403, reshaping them into a long format for comparative analysis, and returns this dataset. 

After the function returns, the simple returns, this is converted into cumalative returns. Ultimately, all the cumulative returns for sectors/indices are joined into one using rbind. We are left with two data frames, one for Sector and one for Index
```{r warning=FALSE}
ALSI <- read_rds("data/ALSI.rds")
RebDays <- read_rds("data/Rebalance_days.rds")
ZAR <- read_rds("data/Monthly_zar.rds")


#Calculate the returns for ALSI and SWIX, grouped by SECTOR
sectors <- ALSI %>% pull(Sector) %>% unique()
sectorReturns <- list()

for(i in 1:length(sectors)){
    sectorReturns[[i]] <- calc_portfolio_returns(ALSI, chosen_sector = sectors[i]) %>% 
        group_by(Fund) %>% 
        mutate(CumReturn = (cumprod(1 + Returns))) %>% 
        mutate(CumReturn = CumReturn/first(CumReturn)) %>% 
        mutate(Sector = sectors[i])
}
names(sectorReturns) <- sectors

FullSectors_df <- rbind(sectorReturns[[1]],
                         sectorReturns[[2]], 
                         sectorReturns[[3]],
                         sectorReturns[[4]]) %>% arrange(date)

#Calculate the returns for ALSI and SWIX, grouped by INDEX
indices <- ALSI %>%  pull(Index_Name) %>% na.omit(.) %>%  unique()
IndexReturns<- list()

for(i in 1:length(indices)){
    IndexReturns[[i]] <- calc_portfolio_returns(ALSI, chosen_index = indices[i]) %>% 
        group_by(Fund) %>% 
        mutate(CumReturn = (cumprod(1 + Returns))) %>% 
        mutate(CumReturn = CumReturn/first(CumReturn)) %>% 
        mutate(Index = indices[i])
}
# Rename tibbles
names(IndexReturns) <- indices
# Combine Dataframes

FullIndex_df <- rbind(IndexReturns[[1]],
                      IndexReturns[[2]], 
                      IndexReturns[[3]]) %>% arrange(date)

```

Next I plot the cumulative returns for both ALSI and SWIX, grouped first by sectors, then by index. The area under each curve is shaded. The row binded data frames above allow me to now use 'facet_wrap' and instantly have seperate plots for different sectors/indices.

```{r warning = FALSE}
#Create 4 facet wrapped plots of sector cumulative returns
sector_plot <- ggplot(data = FullSectors_df, aes(x = date, y = CumReturn, group = Fund)) +
    geom_line(aes(color = Fund)) +
    geom_ribbon(data = filter(FullSectors_df, Fund == "J203"), aes(ymin = 0, ymax = CumReturn, fill = "J203"), alpha = 0.3) +
    geom_ribbon(data = filter(FullSectors_df, Fund == "J403"), aes(ymin = 0, ymax = CumReturn, fill = "J403"), alpha = 0.3) +
    scale_color_manual(values = c("J203" = "blue", "J403" = "red")) +
    scale_fill_manual(values = c("J203" = "lightblue", "J403" = "pink")) +
    guides(fill = guide_none()) +
    scale_x_date(date_labels = "%Y-%m-%d") +
    labs(title = "Cumulative Returns by Sector", x = "Date", y = "Cumulative Return") +
    theme_minimal()+
    facet_wrap(~ Sector, scales = "free_y", ncol = 1)+
    theme(legend.position = "bottom")

finplot(sector_plot)

#Create 4 facet wrapped plots of index cumulative returns
index_plot <- ggplot(data = FullIndex_df, aes(x = date, y = CumReturn, group = Fund)) +
    geom_line(aes(color = Fund)) +
    geom_ribbon(data = filter(FullIndex_df, Fund == "J203"), aes(ymin = 0, ymax = CumReturn, fill = "J203"), alpha = 0.3) +
    geom_ribbon(data = filter(FullIndex_df, Fund == "J403"), aes(ymin = 0, ymax = CumReturn, fill = "J403"), alpha = 0.3) +
    scale_color_manual(values = c("J203" = "blue", "J403" = "red")) +
    scale_fill_manual(values = c("J203" = "lightblue", "J403" = "pink")) +
    guides(fill = guide_none()) +
    scale_x_date(date_labels = "%Y-%m-%d") +
    labs(title = "Cumulative Returns by Index", x = "Date", y = "Cumulative Return") +
    theme_minimal()+
    facet_wrap(~ Index, scales = "free_y", ncol = 1)+
    theme(legend.position = "bottom")

finplot(index_plot)
```

When looking at Sectors, the SWIX(J403) and ALSI(J203) perform nearly identical when looking at the financial and property sector. ALSI outperforms SWIX in the industrial sector, but SWIX does better in resources. When looking at Index, interestingly, the ALSI outperforms SWIX across the board.

The next step is to do stratification analysis with the Rand. This is done in the code chunk below. I first filter currency data to start at the same time as the ALSI and SWIX data, and then calculate the returns. Next I convert to yearly since the data is montly and calculate the 80% and 20% quantiles of the returns and save every year that exceeds those quantiles in lists. I then run the 'calc_portfolio_returns' function again as well as calculate the 1% quantiles for the portfolio returns.

Next I apply a function callled 'Perf_comparisons'. This function takes in the returns as well as the high and low volatility periods and returns a data frame that contains the full standard deviation for the index, as well as its standard deviation during either a high or low volatility period of the Rand. 
```{r warning=FALSE}
ZAR_returns <- ZAR %>% 
    filter(date > '2013-01-02') %>% #Filter for ZAR data after ALSI data start date
    mutate(Return = value/lag(value)-1) %>% 
    filter(date > first(date)) %>% 
    select(c(date, Return))


ZAR_sd <- ZAR_returns %>% 
    mutate(Year = format(date, '%Y')) %>% 
    group_by(Year) %>% 
    summarise(SD = sd(Return)*sqrt(12)) %>% 
  
    mutate(TopQuantile = quantile(SD, 0.8), BotQuantile = quantile(SD, 0.2))

Hi_Vol <- ZAR_sd %>%  filter(SD > TopQuantile) %>% pull(Year)
Low_Vol <- ZAR_sd %>%  filter(SD < BotQuantile) %>% pull(Year)


ReturnsDF <- calc_portfolio_returns(ALSI) %>% 
    mutate(Year= format(date, "%Y")) %>% 
    rename(Tickers = Fund, Return = Returns) %>% 
    group_by(Tickers) %>% 
    mutate(Top = quantile(Return, 0.99), Bot = quantile(Return, 0.01)) %>% 
    mutate(Return = ifelse(Return > Top, Top, 
                         ifelse(Return < Bot, Bot, Return))) %>% ungroup()

perf_hi <- Perf_comparisons(ReturnsDF, Ys = Hi_Vol, Alias = "High_Vol")
perf_lo <- Perf_comparisons(ReturnsDF, Ys = Low_Vol, Alias = "Low_Vol")

perf_hi
perf_lo
```

Ratios above one indicate that that Ticker has a high (above usual) volatility for that given period. In the case of high volatility of the Rand, both the ALSI and SWIX show above usual volatility. Interestingly, ALSI also shows above usual volatility during periods of low Rand volatility.

I now continue on to construct a capped portfolio for both the ALSI and SWIX. The first step is to load the rebalance days from the data. Next I create a function, called 'calculate_capped_port' that applies the 'Proportional_Cap_Foo' function (straight from the practical) to this data. It then gets the returns and weights in xts form which allows me to construct a portfolio with the 'rmsfuns::Safe_Return.portfolio' function. For both ALSI and SWIX, I filter for days that are rebalance days and reformat the data frame for the function. I then calculate portfolios for a 10% and 5% cap for ALSI and SWIX. Lastly, I calculate the 3 year rolling annualised return for each portfolio. Lastly, i join these rolling returns into 2 data frames, a 5% cap and a 10% cap.
```{r}
RebDays <- RebDays %>% 
    filter(Date_Type == 'Reb Trade Day')

calculate_capped_port <- function(dataframe, W_cap, ColName) {

    Capped_df <- dataframe %>% 
        group_split(RebalanceTime) %>% 
        map_df(~Proportional_Cap_Foo(., W_Cap = W_cap)) %>% 
        select(-RebalanceTime)

    df_weights <- Capped_df %>% 
        tbl_xts(cols_to_xts = weight, spread_by = Tickers)

    df_returns <- dataframe %>% 
        filter(Tickers %in% unique(Capped_df$Tickers)) %>% 
        tbl_xts(cols_to_xts = Return, spread_by = Tickers)

    df_weights[is.na(df_weights)] <- 0
    df_returns[is.na(df_returns)] <- 0

    df_capped <- rmsfuns::Safe_Return.portfolio(R = df_returns, 
                                                  weights = df_weights, 
                                                  lag_weights = TRUE) %>% 
        xts_tbl()
    
    colnames(df_capped) <- c('date', ColName)

    return(df_capped)
}

### ALSI
#Get data ready  for function
rebalance_ALSI <- ALSI %>% 
    filter(date %in% RebDays$date) %>% 
    mutate(RebalanceTime = format(date, "%Y%B")) %>% 
    select(date, Tickers, Return, J203, RebalanceTime) %>% 
    rename(weight = J203) %>% 
    mutate(weight = coalesce(weight , 0))

ALSI_capped_10 <- calculate_capped_port(rebalance_ALSI, 0.1, 'ALSI_10')
ALSI_capped_5 <- calculate_capped_port(rebalance_ALSI, 0.05, "ALSI_5")

ALSI_10_rollrets <- ALSI_capped_10 %>% 
    mutate(RollRets_ALSI_10 = RcppRoll::roll_prod(1 + ALSI_10, 4, fill = NA, align = "right")^(4/12)-1) %>% 
    group_by(date) %>% 
    filter(any(!is.na(RollRets_ALSI_10))) %>% 
    ungroup() %>% 
    select(date, RollRets_ALSI_10)

ALSI_5_rollrets <- ALSI_capped_5 %>% 
    mutate(RollRets_ALSI_5 = RcppRoll::roll_prod(1 + ALSI_5, 4, fill = NA, align = "right")^(4/12)-1) %>% 
    group_by(date) %>% 
    filter(any(!is.na(RollRets_ALSI_5))) %>% 
    ungroup() %>% 
    select(date, RollRets_ALSI_5)



###SWIX
rebalance_SWIX <- ALSI %>% 
    filter(date %in% RebDays$date) %>% 
    mutate(RebalanceTime = format(date, "%Y%B")) %>% 
    select(date, Tickers, Return, J403, RebalanceTime) %>% 
    rename(weight = J403) %>% 
    mutate(weight = coalesce(weight , 0))
  
SWIX_capped_10 <- calculate_capped_port(rebalance_SWIX, 0.1, 'SWIX_10')
SWIX_capped_5 <- calculate_capped_port(rebalance_SWIX, 0.05, 'SWIX_5')

SWIX_10_rollrets <- SWIX_capped_10 %>% 
    mutate(RollRets_SWIX_10 = RcppRoll::roll_prod(1 + SWIX_10, 4, fill = NA, align = "right")^(4/12)-1) %>% 
    group_by(date) %>% 
    filter(any(!is.na(RollRets_SWIX_10))) %>% 
    ungroup() %>% 
    select(date, RollRets_SWIX_10)

SWIX_5_rollrets <- SWIX_capped_5 %>% 
    mutate(RollRets_SWIX_5 = RcppRoll::roll_prod(1 + SWIX_5, 4, fill = NA, align = "right")^(4/12)-1) %>% 
    group_by(date) %>% 
    filter(any(!is.na(RollRets_SWIX_5))) %>% 
    ungroup() %>% 
    select(date, RollRets_SWIX_5)


cap5_df <- left_join(ALSI_5_rollrets, SWIX_5_rollrets, by = 'date')
plotdf_5 <- cap5_df %>% 
    pivot_longer(-date,
                 names_to = 'Fund',
                 values_to = 'RollRet')

cap10_df <- left_join(ALSI_10_rollrets, SWIX_10_rollrets, by = 'date')
plotdf_10 <- cap10_df %>% 
    pivot_longer(-date,
                 names_to = 'Fund',
                 values_to = 'RollRet')
```

Lastly, I plot the rolling returns grouped by the the portfolio cap.

```{r warning=FALSE}
g <- plotdf_5 %>% 
    ggplot() + 
    geom_line(aes(date, RollRet, color = Fund), alpha = 0.7, size = 1.25) + 
    labs(title = "Rolling 3 Year Annualized Returns SWIX vs ALSI (5% cap)", 
        subtitle = "", x = "", y = "Rolling 3 year Returns (Ann.)", 
        caption = "") + 
    theme_fmx(title.size = ggpts(30), subtitle.size = ggpts(5), caption.size = ggpts(25), CustomCaption = F) + 
    fmx_cols()+
    scale_fill_brewer(palette = "Accent") 

finplot(g, x.date.dist = "1 year", x.date.type = "%Y", x.vert = T, 
    y.pct = T, y.pct_acc = 1)

g2 <- plotdf_10 %>% 
    ggplot() + 
    geom_line(aes(date, RollRet, color = Fund), alpha = 0.7, size = 1.25) + 
    labs(title = "Rolling 3 Year Annualized Returns SWIX vs ALSI (10% cap)", 
        subtitle = "", x = "", y = "Rolling 3 year Returns (Ann.)", 
        caption = "") + 
    theme_fmx(title.size = ggpts(30), subtitle.size = ggpts(5), caption.size = ggpts(25), CustomCaption = F) + 
    fmx_cols()+
    scale_fill_brewer(palette = "Accent") 

finplot(g2, x.date.dist = "1 year", x.date.type = "%Y", x.vert = T, 
    y.pct = T, y.pct_acc = 1)
```

The ALSI and the SWIX's rolling 3 year annualized returns closely follow each other. There is not much difference between the two for both a 5% and 10% cap. Interestingly, when looking at the period leading up to and straight after Covid-19, ALSI outperforms SWIX during the periods of high return, but under-performs during the downswing. This gap widens when looking at 10% capped portfolios.

# Question 4

This question looks at several funds, their returns and the flows of capital to and from those funds. They key question is then, do short term successes have an effect of flow of funds? To answer this I calculate correlation between short term returns and flows. I then investigate whether their exists any notable correlation between these correlations, fund size and average returns.

I load the data and write a small function to calculate rolling returns for any amount of months I want, using 'rollapply' function from the 'zoo' package. I then calculate the rolling returns for each fund for 6 months, 1 year and 3 years. I then join this with the flows data.

```{r}
Flows <- read_rds("data/ASISA_Flows.rds")
Rets <- read_rds("data/ASISA_Rets.rds")

calculate_period_return <- function(data, months) {
  data %>% 
    group_by(Fund) %>%
    arrange(Fund, date) %>%
    mutate(RollingReturn = rollapply((1 + Returns), width = months, prod, fill = NA, align = 'right')) %>%
    mutate(RollingReturn = RollingReturn - 1) # Subtracting 1 to get the actual return
}

# Calculate 6-month, 1-year, and 3-year returns
rets_6m <- calculate_period_return(Rets, 6) %>% 
    select(date, Fund, RollingReturn)
rets_1y <- calculate_period_return(Rets, 12) %>% 
    select(date, Fund, RollingReturn)
rets_3y <- calculate_period_return(Rets, 36) %>% 
    select(date, Fund, RollingReturn)

#Join with flows
joined_6m <- inner_join(rets_6m, Flows, by = c('date', 'Fund')) %>% 
    filter(!is.na(RollingReturn) & !is.na(Flows)) %>% 
    select(c(-Index, -FoF))
    
joined_1y <- inner_join(rets_1y, Flows, by = c('date', 'Fund')) %>% 
    filter(!is.na(RollingReturn) & !is.na(Flows)) %>% 
    select(c(-Index, -FoF))

joined_3y <- inner_join(rets_3y, Flows, by = c('date', 'Fund')) %>% 
    filter(!is.na(RollingReturn) & !is.na(Flows)) %>% 
    select(c(-Index, -FoF))


```

Next I calculate the correlation between the rolling returns and the next months flows for each fund in the code below. This is then joined, averaged, and tabled.

```{r warning=FALSE}
cor_6m <- joined_6m %>%
  group_by(Fund) %>%
  mutate(next_Flows = lead(Flows)) %>%
  filter(!is.na(next_Flows)) %>%
  summarise(correlation = cor(RollingReturn, next_Flows, use = "complete.obs"))

cor_1y <- joined_1y %>%
  group_by(Fund) %>%
  mutate(next_Flows = lead(Flows)) %>%
  filter(!is.na(next_Flows)) %>%
  summarise(correlation = cor(RollingReturn, next_Flows, use = "complete.obs"))

cor_3y <- joined_3y %>%
  group_by(Fund) %>%
  mutate(next_Flows = lead(Flows)) %>%
  filter(!is.na(next_Flows)) %>%
  summarise(correlation = cor(RollingReturn, next_Flows, use = "complete.obs"))

joined_cor <- left_join(cor_6m, cor_1y, by = "Fund") %>% 
    left_join(cor_3y, by = 'Fund')
colnames(joined_cor) <- c("Fund", "cor_6m", "cor_1y", "cor_3y")

avg_6m_cor <- mean(cor_6m$correlation, na.rm = T)
avg_1y_cor <- mean(cor_1y$correlation, na.rm = T)
avg_3y_cor <- mean(cor_3y$correlation, na.rm = T)


cor_table <- data.frame(avg_6m_cor, avg_1y_cor, avg_3y_cor)
colnames(cor_table) <- c('6 Month', '1 Year', '3 Year')

result <- kable(cor_table)
result
```

The table above shows the average correlation between rolling returns and flows to a fund. As can be seen, very little correlation exists. The highest correlation is an 8% correlation for the 1 year rolling returns. This early result indicates that investors know that past wins do not mean future wins and as such don't flock to a fund that has been performing well recently.

Next I calculate the total flows and the average returns for each fund and join this with the correlation dataframe.

```{r}
flow_df <- Flows %>% 
    group_by(Fund) %>% 
    summarise(TotalFlows = sum(Flows, na.rm = T))

return_df <- Rets %>% 
    group_by(Fund) %>% 
    summarise(avgReturn = mean(Returns, na.rm = T))

joined_df <- left_join(flow_df, return_df, by = 'Fund') %>% 
    left_join(joined_cor, by = "Fund")
```


```{r warning=FALSE}
g1 <- joined_df %>% 
    arrange(desc(avgReturn)) %>% 
    slice(1:50) %>%  #Only use top 50 funds to make plot more understandable
    ggplot(aes(x=TotalFlows, y=cor_6m, size=avgReturn))+
    geom_point(alpha=0.7, color = "steelblue")+
    ylim(c(-0.8, 1.3))+
    scale_size(range = c(3, 15), name="Average Returns")+
    labs(title = "6 Month Correlation vs Total Flows",
         x = 'Total Flows',
         y = '1 Year Correlation')+ 
    theme_fmx(title.size = ggpts(30), subtitle.size = ggpts(5), caption.size = ggpts(25), CustomCaption = F) + 
    fmx_cols()

finplot(g1)
```

```{r warning=FALSE}
g2 <- joined_df %>% 
    arrange(desc(avgReturn)) %>% 
    slice(1:50) %>% 
    ggplot(aes(x=TotalFlows, y=cor_1y, size=avgReturn))+
    geom_point(alpha=0.7, color = "lightsalmon3")+
    scale_size(range = c(3, 15), name="Average Returns")+
    labs(title = "1 Year Correlation vs Total Flows",
         x = 'Total Flows',
         y = '1 Year Correlation')+ 
    theme_fmx(title.size = ggpts(30), subtitle.size = ggpts(5), caption.size = ggpts(25), CustomCaption = F) + 
    fmx_cols()
finplot(g2)
```

```{r warning=FALSE}
g3 <- joined_df %>% 
    arrange(desc(avgReturn)) %>% 
    slice(1:50) %>% 
    ggplot(aes(x=TotalFlows, y=cor_3y, size=avgReturn))+
    geom_point(alpha=0.7, color = "orange")+
    scale_size(range = c(3, 15), name="Average Returns")+
    labs(title = "3 Year Correlation vs Total Flows",
         x = 'Total Flows',
         y = '1 Year Correlation')+ 
    theme_fmx(title.size = ggpts(30), subtitle.size = ggpts(5), caption.size = ggpts(25), CustomCaption = F) + 
    fmx_cols()
finplot(g3)
```


The plots above shows scatter plots between total flows for each fund and the correlation of returns and flows. The size of the points are determined by the average returns for that fund. It is clear that no correlation exists. This means that the total flows a fund receive does not change the perception of investors. They still hold the belief that past successes don't predict future winners. The average returns of a fund also don't seem to have an effect. It it did we would see all the large points gather at either the top or bottom of the graph. This result holds for the 6 month, 1 year and 3 year correlations.
