---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Question 4: Flow Analysis"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Ruan Geldenhuys"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch University, Stellenbosch, South Africa" # First Author's Affiliation
Email1: "22550801\\@sun.ac.za" # First Author's Email address

# Author2: "John Smith"
# Ref2: "Some other Institution, Cape Town, South Africa"
# Email2: "John\\@gmail.com"
# CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.
# 
# Author3: "John Doe"
# Email3: "Joe\\@gmail.com"
# 
# CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
# keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
# JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
# BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
# addtoprule: TRUE
# addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
#Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
#abstract: |
  
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf.
# These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!

# Lets load in example data, and see how this can be stored and later called from your 'data' folder.
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
library(dplyr)
library(lubridate)
library(zoo)
Flows <- read_rds("data/ASISA_Flows.rds")
Rets <- read_rds("data/ASISA_Rets.rds")


```

# Introduction

This question looks at several funds, their returns and the flows of capital to and from those funds. They key question is then, do short term successes have an effect of flow of funds? To answer this I calculate correlation between short term returns and flows. I then investigate whether their exists any notable correlation between these correlations, fund size and average returns.



```{r}
library(dplyr)
library(lubridate)
library(zoo)


Rets <- Rets %>% 
  mutate(date = as.Date(date, format = "%Y-%m-%d"))

calculate_period_return <- function(data, months) {
  data %>% 
    group_by(Fund) %>%
    arrange(Fund, date) %>%
    mutate(RollingReturn = rollapply((1 + Returns), width = months, prod, fill = NA, align = 'right')) %>%
    mutate(RollingReturn = RollingReturn - 1) # Subtracting 1 to get the actual return
}

# Calculate 6-month, 1-year, and 3-year returns
rets_6m <- calculate_period_return(Rets, 6) %>% 
    select(date, Fund, RollingReturn)
rets_1y <- calculate_period_return(Rets, 12) %>% 
    select(date, Fund, RollingReturn)
rets_3y <- calculate_period_return(Rets, 36) %>% 
    select(date, Fund, RollingReturn)



```


```{r warning=FALSE}
library(dplyr)
library(zoo)
library(kableExtra)


joined_6m <- inner_join(rets_6m, Flows, by = c('date', 'Fund')) %>% 
    filter(!is.na(RollingReturn) & !is.na(Flows)) %>% 
    select(c(-Index, -FoF))
    
joined_1y <- inner_join(rets_1y, Flows, by = c('date', 'Fund')) %>% 
    filter(!is.na(RollingReturn) & !is.na(Flows)) %>% 
    select(c(-Index, -FoF))

joined_3y <- inner_join(rets_3y, Flows, by = c('date', 'Fund')) %>% 
    filter(!is.na(RollingReturn) & !is.na(Flows)) %>% 
    select(c(-Index, -FoF))



    
cor_6m <- joined_6m %>%
  group_by(Fund) %>%
  mutate(next_Flows = lead(Flows)) %>%
  filter(!is.na(next_Flows)) %>%
  summarise(correlation = cor(RollingReturn, next_Flows, use = "complete.obs"))

cor_1y <- joined_1y %>%
  group_by(Fund) %>%
  mutate(next_Flows = lead(Flows)) %>%
  filter(!is.na(next_Flows)) %>%
  summarise(correlation = cor(RollingReturn, next_Flows, use = "complete.obs"))

cor_3y <- joined_3y %>%
  group_by(Fund) %>%
  mutate(next_Flows = lead(Flows)) %>%
  filter(!is.na(next_Flows)) %>%
  summarise(correlation = cor(RollingReturn, next_Flows, use = "complete.obs"))

joined_cor <- left_join(cor_6m, cor_1y, by = "Fund") %>% 
    left_join(cor_3y, by = 'Fund')
colnames(joined_cor) <- c("Fund", "cor_6m", "cor_1y", "cor_3y")

avg_6m_cor <- mean(cor_6m$correlation, na.rm = T)
avg_1y_cor <- mean(cor_1y$correlation, na.rm = T)
avg_3y_cor <- mean(cor_3y$correlation, na.rm = T)


cor_table <- data.frame(avg_6m_cor, avg_1y_cor, avg_3y_cor)
colnames(cor_table) <- c('6 Month', '1 Year', '3 Year')

result <- kable(cor_table)
result
```

The table above shows the average correlation between rolling returns and flows to a fund. As can be seen, very little correlation exists. The highest correlation is an 8% correlation for the 1 year rolling returns. This early result indicates that investors know that past wins do not mean future wins and as such don't flock to a fund that has been performing well recently.

```{r Figure1,  warning =  FALSE, fig.align = 'center', fig.cap = "6 Month Correlation vs Total Flows", fig.ext = 'png', fig.height = 3, fig.width = 6}
library(fmxdat)
flow_df <- Flows %>% 
    group_by(Fund) %>% 
    summarise(TotalFlows = sum(Flows, na.rm = T))

return_df <- Rets %>% 
    group_by(Fund) %>% 
    summarise(avgReturn = mean(Returns, na.rm = T))

joined_df <- left_join(flow_df, return_df, by = 'Fund') %>% 
    left_join(joined_cor, by = "Fund")

g1 <- joined_df %>% 
    arrange(desc(avgReturn)) %>% 
    slice(1:50) %>%  #Only use top 50 funds to make plot more understandable
    ggplot(aes(x=TotalFlows, y=cor_6m, size=avgReturn))+
    geom_point(alpha=0.7, color = "steelblue")+
    ylim(c(-0.8, 1.3))+
    scale_size(range = c(3, 15), name="Average Returns")+
    labs(title = "6 Month Correlation vs Total Flows",
         x = 'Total Flows',
         y = '1 Year Correlation')+ 
    theme_fmx(title.size = ggpts(30), subtitle.size = ggpts(5), caption.size = ggpts(25), CustomCaption = F) + 
    fmx_cols()

finplot(g1)





```

```{r Figure2,  warning =  FALSE, fig.align = 'center', fig.cap = "1 Year Correlation vs Total Flows", fig.ext = 'png', fig.height = 3, fig.width = 6}
g2 <- joined_df %>% 
    arrange(desc(avgReturn)) %>% 
    slice(1:50) %>% 
    ggplot(aes(x=TotalFlows, y=cor_1y, size=avgReturn))+
    geom_point(alpha=0.7, color = "lightsalmon3")+
    scale_size(range = c(3, 15), name="Average Returns")+
    labs(title = "1 Year Correlation vs Total Flows",
         x = 'Total Flows',
         y = '1 Year Correlation')+ 
    theme_fmx(title.size = ggpts(30), subtitle.size = ggpts(5), caption.size = ggpts(25), CustomCaption = F) + 
    fmx_cols()
finplot(g2)
```

```{r Figure3,  warning =  FALSE, fig.align = 'center', fig.cap = "3 Year Correlation vs Total Flows", fig.ext = 'png', fig.height = 3, fig.width = 6}
g3 <- joined_df %>% 
    arrange(desc(avgReturn)) %>% 
    slice(1:50) %>% 
    ggplot(aes(x=TotalFlows, y=cor_3y, size=avgReturn))+
    geom_point(alpha=0.7, color = "orange")+
    scale_size(range = c(3, 15), name="Average Returns")+
    labs(title = "3 Year Correlation vs Total Flows",
         x = 'Total Flows',
         y = '1 Year Correlation')+ 
    theme_fmx(title.size = ggpts(30), subtitle.size = ggpts(5), caption.size = ggpts(25), CustomCaption = F) + 
    fmx_cols()
finplot(g3)

```

The plots above shows scatter plots between total flows for each fund and the correlation of returns and flows. The size of the points are determined by the average returns for that fund. It is clear that no correlation exists. This means that the total flows a fund receive does not change the perception of investors. They still hold the belief that past successes don't predict future winners. The average returns of a fund also don't seem to have an effect. It it did we would see all the large points gather at either the top or bottom of the graph. This result holds for the 6 month, 1 year and 3 year correlations.

