---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Question 2: Currency Hedging Analysis"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Ruan Geldenhuys"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch University, Stellenbosch, South Africa" # First Author's Affiliation
Email1: "22550801\\@sun.ac.za" # First Author's Email address

# Author2: "John Smith"
# Ref2: "Some other Institution, Cape Town, South Africa"
# Email2: "John\\@gmail.com"
# CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.
# 
# Author3: "John Doe"
# Email3: "Joe\\@gmail.com"
# 
# CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
# keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
# JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
# BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
# addtoprule: TRUE
# addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
#Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
#abstract: |
  
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf.
# These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!

# Lets load in example data, and see how this can be stored and later called from your 'data' folder.
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
library(ggExtra)
library(PerformanceAnalytics)
library(tbl2xts)
library(fmxdat)


```

# Introduction

This question focuses on a recent article that discusses the misconception surrounding hedging out the Rand. By replicating the figure within that article I therefore show that hedging very rarely has beneficial effects. I then go further by showing how the volatility of a portfolio can actually be increased through hedging


```{r Figure1,  warning =  FALSE, fig.align = 'center', fig.cap = "Replicated Figure", fig.ext = 'png', fig.height = 3.5, fig.width = 7}
Indexes <- read_rds("data/Cncy_Hedge_Assets.rds")
ZAR <- read_rds("data/Monthly_zar.rds")

weights <- c(0.18, 0.12, 0.42, 0.28)

Portfolio <- Indexes %>% 
    mutate(MSCI_ACWI = MSCI_ACWI * 0.18) %>% 
    mutate(Bbg_Agg = Bbg_Agg * 0.12) %>% 
    mutate(J433 = J433 * 0.42) %>% 
    mutate(ALBI = ALBI * 0.28)

PortfolioReturns <- Portfolio %>%
    rowwise() %>%
    mutate(sum = sum(c_across(-date), na.rm = TRUE)) %>%
    select(date, sum)


ZAR_Returns <- ZAR %>% 
    mutate(Return = value/lag(value) - 1) %>% 
    select(c('date','Return'))

joinedDF <- left_join(PortfolioReturns, ZAR_Returns, by = 'date')
colnames(joinedDF)<- c('date','Portfolio_Returns','USDZAR_Returns')

#Calculate Quadrant Percentages
cleanJoined <- na.omit(joinedDF)
total_count <- nrow(cleanJoined)

bothPositive_count <- sum(cleanJoined$Portfolio_Returns > 0 & cleanJoined$USDZAR_Returns > 0, na.rm = TRUE)
percentage_positive <- round((bothPositive_count / total_count) * 100, 1)
percentage_positive <- paste0(percentage_positive, "%")

bothNegative_count <- sum(cleanJoined$Portfolio_Returns < 0 & cleanJoined$USDZAR_Returns < 0, na.rm = TRUE)
percentage_negative <- round((bothNegative_count / total_count) * 100, 1)
percentage_negative <- paste0(percentage_negative, "%")

PosAndNeg_count <- sum(cleanJoined$Portfolio_Returns > 0 & cleanJoined$USDZAR_Returns < 0, na.rm = TRUE)
percentage_PosAndNeg <- round((PosAndNeg_count / total_count) * 100, 1)
percentage_PosAndNeg <- paste0(percentage_PosAndNeg, "%")

NegAndPos_count <- sum(cleanJoined$Portfolio_Returns < 0 & cleanJoined$USDZAR_Returns > 0, na.rm = TRUE)
percentage_NegAndPos <- round((NegAndPos_count / total_count) * 100, 1)
percentage_NegAndPos <- paste0(percentage_NegAndPos, "%")

TopLeft_hedging <- sum(cleanJoined$Portfolio_Returns > 0 & cleanJoined$USDZAR_Returns < -0.025, na.rm = TRUE)
percentage_TopLeft_hedge <- round((TopLeft_hedging / total_count) * 100, 1)
percentage_TopLeft_hedge <- paste0(percentage_TopLeft_hedge, "%")

BotLeft_hedging <- sum(cleanJoined$Portfolio_Returns < 0 & cleanJoined$USDZAR_Returns < -0.025, na.rm = TRUE)
percentage_BotLeft_hedge <- round((BotLeft_hedging / total_count) * 100, 1)
percentage_BotLeft_hedge <- paste0(percentage_BotLeft_hedge, "%")



#Plotting this monster
p<- ggplot(joinedDF, aes(x = USDZAR_Returns, y = Portfolio_Returns)) +
    geom_point() +  # Add points for scatterplot
    geom_smooth(method = "lm", color = "blue") +  # Add linear regression line
    ylim(-0.2,0.2)+
    xlim(-0.2,0.2)+
    
    geom_vline(xintercept = 0, linetype = "solid") +  # Add vertical line at zero
    geom_hline(yintercept = 0, linetype = "solid") +  # Add horizontal line at zero
    
    geom_vline(xintercept = -0.025, linetype = "longdash")+
    
    annotate("rect", xmin = -Inf, xmax = 0, ymin = 0, ymax = Inf, fill = "green", alpha = 0.2) + # Top-left quadrant
    annotate("rect", xmin = 0, xmax = Inf, ymin = 0, ymax = Inf, fill = "red", alpha = 0.2) +   # Top-right quadrant
    annotate("rect", xmin = -Inf, xmax = 0, ymin = -Inf, ymax = 0, fill = "pink", alpha = 0.2) + # Bottom-left quadrant
    annotate("rect", xmin = 0, xmax = Inf, ymin = -Inf, ymax = 0, fill = "blue", alpha = 0.2) + # Bottom-right quadrant
    
    labs(title = "ZAR/USD Returns vs. Portfolio Returns",
       x = "USD/ZAR Returns",
       y = "Portfolio Returns") +
    theme_minimal()+
    
    geom_label(aes(x = -Inf, y = Inf, label = percentage_PosAndNeg), 
               hjust = 0, vjust = 1, size = 4, color = "black", fill = "white") +  # Top-left quadrant
    geom_label(aes(x = Inf, y = Inf, label = percentage_positive), 
               hjust = 1, vjust = 1, size = 4, color = "black", fill = "white") +  # Top-right quadrant
    geom_label(aes(x = -Inf, y = -Inf, label = percentage_negative), 
               hjust = 0, vjust = 0, size = 4, color = "black", fill = "white") +  # Bottom-left quadrant
    geom_label(aes(x = Inf, y = -Inf, label = percentage_NegAndPos), 
               hjust = 1, vjust = 0, size = 4, color = "black", fill = "white") +  # Bottom-right quadrant
    
    geom_label(aes(x = -0.025, y = Inf, label = percentage_TopLeft_hedge), 
               hjust = 1, vjust = 1, size = 4, color = "black", fill = "white") +
    geom_label(aes(x = -0.025, y = -Inf, label = percentage_BotLeft_hedge), 
               hjust = 1, vjust = 0, size = 4, color = "black", fill = "white") +
    
    geom_label(aes(x=-0.1, y=0.1, label = "Hedge works but \n amplifies volatility"),
               hjust = 0.5, vjust = 0.5, size = 4, color = "black", fill = "white") +
    geom_label(aes(x=0.1, y=0.1, label = "Hedge throws away returns"),
               hjust = 0.5, vjust = 0.5, size = 4, color = "black", fill = "white") +
    geom_label(aes(x=-0.1, y=-0.1, label = "Best case for Hedge"),
               hjust = 0.5, vjust = 0.5, size = 4, color = "black", fill = "white") +
    geom_label(aes(x=0.1, y=-0.1, label = "Hedge removes currency \n cushion"),
               hjust = 0.5, vjust = 0.5, size = 4, color = "black", fill = "white") +
    
     theme(panel.border = element_rect(colour = "black", fill=NA, size=1))



p <- ggMarginal(p, type = "density", fill = 'darkgreen', color = NA)

p




```


The graph interpretation can be done as follows. If we find ourselves in the top left quadrant, it means portfolio returns are positive and USD/ZAR returns are negative. In this case a hedge will work but it increases volatility, a less than desirable outcome. This occurs for 46.2% of observations and 30.1% if we apply a fee. The bottom left quadrant shows when portfolio and USD/ZAR returns are negative. This is the best case scenario for a hedge, yet it only occurs 7.5% of the time. 2.2% when applying a fee. The top right and bottom right quadrant shows outright negative positions to be in when hedging against the Rand. These occur 19.9% and 25.8% respectively.


```{r Figure2,  warning =  FALSE, fig.align = 'center', fig.cap = "Rolling 3 Year Annualized SD", fig.ext = 'png', fig.height = 4, fig.width = 6}
colnames(PortfolioReturns) <- c('date', 'Return')

weights <- c(0.18, 0.12, 0.42, 0.28)

#Calculate returns and then Rolling SD of a hedged portfolio
#i.e. you get the returns
port1 <- Return.portfolio(Indexes, weights = weights, rebalance_on = "quarters")

hedged_annualised_return <- Return.annualized(port1, scale = 12)
hedged_annualised_std <- StdDev.annualized(port1, scale = 12) 

port_hedged_df <- xts_tbl(port1) %>% 
    mutate(RollSD = RcppRoll::roll_sd(1 + portfolio.returns, 36, fill = NA, align = "right") * 
    sqrt(12)) %>% 
    filter(!is.na(RollSD))


#Calculate returns and then rolling SD of an unhedged portfolio
port_unhedged_df <- xts_tbl(port1) %>% 
    left_join(ZAR_Returns, by = 'date')
colnames(port_unhedged_df) <- c('date', 'portfolioReturns', 'currencyReturns')

port_unhedged_df <- port_unhedged_df %>% 
    mutate(Unhedged_Returns = (0.95*portfolioReturns) + (0.5*currencyReturns)) %>% 
    select(c(date, Unhedged_Returns)) %>% 
    na.omit()

unhedged_annualised_return <- Return.annualized(port_unhedged_df, scale = 12)
unhedged_annualised_std <- StdDev.annualized(port_unhedged_df, scale = 12)

port_unhedged_df <- port_unhedged_df %>% 
    mutate(RollSD_unhedged = RcppRoll::roll_sd(1 + Unhedged_Returns, 36, fill = NA, align = "right") * 
    sqrt(12)) %>% 
    filter(!is.na(RollSD_unhedged))   

#Join the two portfolio SD's and the convert to long
joined_portfolios <-
    left_join(port_hedged_df, port_unhedged_df, by = 'date') %>% 
    select(c(date, RollSD, RollSD_unhedged)) %>% 
    filter(!is.na(RollSD_unhedged)) %>% 
    pivot_longer(-date,
                 names_to = 'portfolio',
                 values_to = 'RollSD')


#Plot
g <- joined_portfolios %>% 
    ggplot() + 
    geom_line(aes(date, RollSD, color = portfolio), alpha = 0.7, size = 1.25) + 
        
    labs(title = "Rolling 3 Year Annualized SD", 
        subtitle = "", x = "", y = "Rolling 3 year SD (Ann.)", 
        caption = "") + theme_fmx(title.size = ggpts(30), 
        subtitle.size = ggpts(5), caption.size = ggpts(25), CustomCaption = F) + 
        
    fmx_cols()
    
    finplot(g, x.date.dist = "1 year", x.date.type = "%Y", x.vert = T, 
        y.pct = T, y.pct_acc = 1)

```

The graph displayed above illustrates the rolling three-year annualized standard deviation for both a hedged and an unhedged portfolio. What stands out is that, for most of the observed period, the unhedged portfolio exhibits lower volatility. This observation reinforces the points made earlier, demonstrating that hedging against the Rand actually elevates the portfolio's overall volatility. This outcome contradicts the primary goal of hedging against the Rand, which is to reduce volatility.



