---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Question 1: Systematic AI Fund"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Ruan Geldenhuys"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch University, Stellenbosch, South Africa" # First Author's Affiliation
Email1: "22550801\\@sun.ac.za" # First Author's Email address

# Author2: "John Smith"
# Ref2: "Some other Institution, Cape Town, South Africa"
# Email2: "John\\@gmail.com"
# CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.
# 
# Author3: "John Doe"
# Email3: "Joe\\@gmail.com"
# 
# CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
# keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
# JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
# BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
# addtoprule: TRUE
# addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
#Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
#abstract: |
  
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')

if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
library(RcppRoll)
library(fmxdat)
library(RColorBrewer)
library(tbl2xts)
library(PerformanceAnalytics)




```

# Question 1

This question investigates returns from an AI Fund, a market benchmark (CAPPED SWIX) and active managers. It aims to shed light on the differences and similarity in return structures through various visualisation techniques. I find that the benchmark often outperforms both the AI Fund and the active managers.

```{r include = FALSE}
#Load and prep the data
ASISA <- read_rds("data/ASISA_Rets.rds")
BM <- read_rds("data/Capped_SWIX.rds")
AI_Fund <- read_rds("data/AI_Max_Fund.rds")
```

```{r Figure1,  warning =  FALSE, fig.align = 'center', fig.cap = "AI vs Active Manager Returns", fig.ext = 'png', fig.height = 3, fig.width = 6}
#Calculate the average ASISA return across all funds for a given month
ASISA_average <- ASISA %>% 
    group_by(date) %>% 
    summarise(meanReturn = mean(Returns))

#Calculate the average BM return across all indices for a given month
BM_average <- BM %>% 
    group_by(date) %>% 
    summarise(meanReturn = mean(Returns))


#Get all the returns in the same dataframe
returnDF <- left_join(ASISA_average, BM_average, by = 'date') %>% 
    left_join(AI_Fund, by='date')
colnames(returnDF)<- c('date','Active', 'Benchmark', 'AI')

returnDF <- returnDF %>% 
    pivot_longer(cols = c(-date),
                 names_to = "Fund",
                 values_to = "Returns")

xts.returnDF <- returnDF %>% 
    tbl_xts(tblData = ., cols_to_xts = Returns, spread_by = Fund)

AIvsActive <- chart.Scatter(x = xts.returnDF$AI, 
                            y = xts.returnDF$Active, 
                            main = "Scatter: AI vs Active Managers", col = "darkred",symbolset = 16, 
                            xlab = "AI", ylab = "Active Managers")

```

The figure above shows a scatter between the returns of active managers and the AI fund. A clear positive correlation is visible. 

```{r Figure2,  warning =  FALSE, fig.align = 'center', fig.cap = "Ai vs Benchmark Returns", fig.ext = 'png', fig.height = 3, fig.width = 6}

AIvsBenchmark <- chart.Scatter(x = xts.returnDF$AI, 
                            y = xts.returnDF$Benchmark, 
                            main = "Scatter: AI vs Benchmark", col = "darkblue",symbolset = 16, 
                            xlab = "Benchmark", ylab = "Benchmark Indices")
```

The figure above now shows a scatter between the returns of the benchmark and the AI fund. An even stronger positive correlation is now visible. This means that for, in both cases, it is rare for the AI fund to show a negative return when the other funds show a positive return and vice-versa.

```{r Figure3,  warning =  FALSE, fig.align = 'center', fig.cap = "Rolling 3 Year Annualized Returns Comparison", fig.ext = 'png', fig.height = 3, fig.width = 6}

##ASISA
#Calculate the annualized three year rolling return
ASISA_RollRet <- ASISA_average %>% 
    mutate(RollRets_ASISA = RcppRoll::roll_prod(1 + meanReturn, 36, fill = NA, align = "right")^(12/36)-1) %>% 
    group_by(date) %>% 
    filter(any(!is.na(RollRets_ASISA))) %>% 
    ungroup() %>%
    select(date, RollRets_ASISA)
    
##BM


##Calculate the annualized three year rolling return
BM_RollRet <- BM_average %>% 
    mutate(RollRets_BM = RcppRoll::roll_prod(1 + meanReturn, 36, fill = NA, align = "right")^(12/36)-1) %>% 
    group_by(date) %>% 
    filter(any(!is.na(RollRets_BM))) %>% 
    ungroup() %>%
    select(date, RollRets_BM)

joinedDF <- left_join(ASISA_RollRet, BM_RollRet, by = "date")

##AIFund
#Calculate the annualized three year rolling return
AI_RollRet <- AI_Fund %>% 
    mutate(RollRets_AI = RcppRoll::roll_prod(1 + AI_Fund, 36, fill = NA, align = "right")^(12/36)-1) %>% 
    group_by(date) %>% 
    filter(any(!is.na(RollRets_AI))) %>% 
    ungroup() %>%
    select(date, RollRets_AI)

joinedDF <- joinedDF %>% 
    left_join(AI_RollRet, by = "date")
colnames(joinedDF) <- c('date','Active','Benchmark','AIFund')


#Plot all 3 rolling returns as line graphs
plottingDF <- joinedDF %>% 
    pivot_longer(cols = c(-date),
                 names_to = "Fund",
                 values_to = "RollingReturns")


g <- plottingDF %>% 
    ggplot() + 
    geom_line(aes(date, RollingReturns, color = Fund), alpha = 0.7, size = 1.25) + 
    labs(title = "Rolling 3 Year Annualized Returns Comparison", 
        subtitle = "", x = "", y = "Rolling 3 year Returns (Ann.)", 
        caption = "") + 
    theme_fmx(title.size = ggpts(30), subtitle.size = ggpts(5), caption.size = ggpts(25), CustomCaption = F) + 
    fmx_cols()+
    scale_fill_brewer(palette = "Accent") 
    

finplot(g, x.date.dist = "1 year", x.date.type = "%Y", x.vert = T, 
    y.pct = T, y.pct_acc = 1)


```

The rolling 3 year annualised returns of all three funds are shown above. All three funds tend to follow each other and show significant downturns during the GFC and Covid-19. Comparatively, the active managers perform the worst, as for large periods of time there rolling returns are substantially lower than the other two funds. The AI fund seems capable of outperforming the benchmark, but never for long periods of time.



```{r Figure4,  warning =  FALSE, fig.align = 'center', fig.cap = "Density of Returns by Funds", fig.ext = 'png', fig.height = 3, fig.width = 6}
returnDF <- returnDF %>% 
    pivot_wider(names_from = Fund, values_from = Returns)
returnDF <- returnDF %>% 
    mutate(Active = Active - 0.025/12) %>%   # Applying a 25bps fee for active and AI fund
    mutate(AI = AI - 0.025/12)
returnDF <- returnDF %>% 
    pivot_longer(cols = c(-date),
                 names_to = "Fund",
                 values_to = "Returns")

# Plot the density functions of each return and add a line for the mean benchmark return
mean_benchmark_returns <- mean(returnDF$Returns[returnDF$Fund == "Benchmark"])
densityPlot <- ggplot(returnDF, aes(x = Returns)) + 
    geom_density(aes(fill = Fund), alpha = 0.8, color = NA) +
    geom_vline(xintercept = mean_benchmark_returns, linetype = "dashed", color = "black") +
    geom_text(aes(x = mean_benchmark_returns, y = 15), label = "Market Benchmark", 
              vjust = -1, hjust = -0.05, color = "black") +
    labs(title = "Density of Returns by Fund",
         x = "Returns",
         y = "Density") +
    theme_minimal() +
    scale_fill_brewer(palette = "Accent") +
    theme(panel.border = element_rect(colour = "black", fill=NA, size=1))

densityPlot

```

The density functions of returns, grouped by fund, show a clear result. Active managers tend to perform below the market benchmark. The marker benchmark here was defined as the average return of the benchmark. The market benchmark and the AI fund perform very similar. The key difference here lies in the flatter distribution of the AI fund, meaning it is less consistent that the market benchmark.


```{r Figure5,  warning =  FALSE, fig.align = 'center', fig.cap = "Comparison of Best and Worst Active funds vs AI Fund", fig.ext = 'png', fig.height = 3, fig.width = 6}
#Find best performing fund
years <- 2003:2023
funds_with_complete_data <- ASISA %>%
  mutate(Year = year(as.Date(date))) %>%
  group_by(Fund) %>%
  summarize(All_Years_Present = all(years %in% unique(Year))) %>%
  filter(All_Years_Present) %>%
  pull(Fund)

BestActive <- ASISA %>%
  filter(Fund %in% funds_with_complete_data) %>%
  group_by(Fund) %>%
  summarise(Average_Return = mean(Returns, na.rm = TRUE)) %>%
  arrange(desc(Average_Return)) %>%
  slice(1)


BestActiveReturns <- ASISA %>% 
    filter(Fund == BestActive$Fund) %>% 
    select(c('date','Returns'))

#Get the rolling returns for best fund
Best_RollRet <- BestActiveReturns %>% 
    mutate(RollRets_best = RcppRoll::roll_prod(1 + Returns, 36, fill = NA, align = "right")^(12/36)-1) %>% 
    group_by(date) %>% 
    filter(any(!is.na(RollRets_best))) %>% 
    ungroup() %>%
    select(date, RollRets_best)



##Find worse performing fund
WorstActive <- ASISA %>%
  filter(Fund %in% funds_with_complete_data) %>%
  group_by(Fund) %>%
  summarise(Average_Return = mean(Returns, na.rm = TRUE)) %>%
  arrange(Average_Return) %>%
  slice(1)


WorstActiveReturns <- ASISA %>% 
    filter(Fund == WorstActive$Fund) %>% 
    select(c('date','Returns'))

Worst_RollRet <- WorstActiveReturns %>% 
    mutate(RollRets_worst = RcppRoll::roll_prod(1 + Returns, 36, fill = NA, align = "right")^(12/36)-1) %>% 
    group_by(date) %>% 
    filter(any(!is.na(RollRets_worst))) %>% 
    ungroup() %>%
    select(date, RollRets_worst)

##Group these funds and AI into one dataset
BestAndWorseDF <- left_join(Best_RollRet, Worst_RollRet, by = 'date') %>% 
    left_join(AI_RollRet, by = 'date')
colnames(BestAndWorseDF) <- c('date', 'BestActive', 'WorstActive', 'AI')


plotDiff <- ggplot(BestAndWorseDF, aes(x = date)) +
  geom_line(aes(y = BestActive, colour = "BestActive")) +
  geom_line(aes(y = WorstActive, colour = "WorstActive")) +
  geom_line(aes(y = AI, colour = "AI")) +
  geom_ribbon(aes(ymin = WorstActive, ymax = BestActive), fill = "grey", alpha = 0.5) +
  labs(title = "Comparison of Best and Worst Active funds vs AI Fund",
       x = "Date",
       y = "Rolling 3 Year Return (Ann.)",
       colour = "Fund") +
  scale_colour_manual(values = c("BestActive" = "blue", "WorstActive" = "red", "AI" = "green")) +
  theme_minimal()+
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1))
plotDiff
```
The figure reveals an interesting story. It appears that the AI fund often lies between the rolling returns of these two funds. Interestingly, for a period in the middle of our sample, the worst performing fund, actually outperformed the best performing fund. This potentially speaks to the shortcomings of average returns as a metric of success. In this period, the AI fund also outperformed both of these fund. The AI fund performs lower than the worst fund at times, but this usually does not last long.

