---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Question 3: Portfolio Construction"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Ruan Geldenhuys"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch University, Stellenbosch, South Africa" # First Author's Affiliation
Email1: "22550801\\@sun.ac.za" # First Author's Email address

# Author2: "John Smith"
# Ref2: "Some other Institution, Cape Town, South Africa"
# Email2: "John\\@gmail.com"
# CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.
# 
# Author3: "John Doe"
# Email3: "Joe\\@gmail.com"
# 
# CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
# keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
# JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
# BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
# addtoprule: TRUE
# addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
#Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
#abstract: |
  
---
<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')

if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
library(pacman)
library(zoo)
library(tbl2xts)

source("code/perf_comparisons.R")
source('code/calculatePortfolioReturns.R')
source('code/PortfolioCapping.R')



```


# Introduction

This questions looks at two different all share indices, namely the ALSI and the SWIX. I compare these two indices by first estimating the cumulative returns for both broken down by sector and index. Next I perform stratification analysis using the ZAR to see how these indices react during currency volatility. Lastly, I compare the differences in returns for different capped portfolios.


```{r}
ALSI <- read_rds("data/ALSI.rds")
RebDays <- read_rds("data/Rebalance_days.rds")
ZAR <- read_rds("data/Monthly_zar.rds")


```


```{r Figure1,  warning =  FALSE, fig.align = 'center', fig.cap = "Cumulative Returns by Sector", fig.ext = 'png', fig.height = 7, fig.width = 7}
library(fmxdat)

sectors <- ALSI %>% pull(Sector) %>% unique()
sectorReturns <- list()

for(i in 1:length(sectors)){
    sectorReturns[[i]] <- calc_portfolio_returns(ALSI, chosen_sector = sectors[i]) %>% 
        group_by(Fund) %>% 
        mutate(CumReturn = (cumprod(1 + Returns))) %>% 
        mutate(CumReturn = CumReturn/first(CumReturn)) %>% 
        mutate(Sector = sectors[i])
}
names(sectorReturns) <- sectors

FullSectors_df <- rbind(sectorReturns[[1]],
                         sectorReturns[[2]], 
                         sectorReturns[[3]],
                         sectorReturns[[4]]) %>% arrange(date)

sector_plot <- ggplot(data = FullSectors_df, aes(x = date, y = CumReturn, group = Fund)) +
    geom_line(aes(color = Fund)) +
    geom_ribbon(data = filter(FullSectors_df, Fund == "J203"), aes(ymin = 0, ymax = CumReturn, fill = "J203"), alpha = 0.3) +
    geom_ribbon(data = filter(FullSectors_df, Fund == "J403"), aes(ymin = 0, ymax = CumReturn, fill = "J403"), alpha = 0.3) +
    scale_color_manual(values = c("J203" = "blue", "J403" = "red")) +
    scale_fill_manual(values = c("J203" = "lightblue", "J403" = "pink")) +
    guides(fill = guide_none()) +
    scale_x_date(date_labels = "%Y-%m-%d") +
    labs(title = "Cumulative Returns by Sector", x = "Date", y = "Cumulative Return") +
    theme_minimal()+
    facet_wrap(~ Sector, scales = "free_y", ncol = 1)+
    theme(legend.position = "bottom")



finplot(sector_plot)



```


```{r Figure1.2,  warning =  FALSE, fig.align = 'center', fig.cap = "Cumulative Returns by Index", fig.ext = 'png', fig.height = 7, fig.width = 7}
# Compare indices for ALSI and SWIX

indices <- ALSI %>%  pull(Index_Name) %>% na.omit(.) %>%  unique()
IndexReturns<- list()

for(i in 1:length(indices)){
    IndexReturns[[i]] <- calc_portfolio_returns(ALSI, chosen_index = indices[i]) %>% 
        group_by(Fund) %>% 
        mutate(CumReturn = (cumprod(1 + Returns))) %>% 
        mutate(CumReturn = CumReturn/first(CumReturn)) %>% 
        mutate(Index = indices[i])
}
# Rename tibbles
names(IndexReturns) <- indices
# Combine Dataframes

FullIndex_df <- rbind(IndexReturns[[1]],
                      IndexReturns[[2]], 
                      IndexReturns[[3]]) %>% arrange(date)
    
index_plot <- ggplot(data = FullIndex_df, aes(x = date, y = CumReturn, group = Fund)) +
    geom_line(aes(color = Fund)) +
    geom_ribbon(data = filter(FullIndex_df, Fund == "J203"), aes(ymin = 0, ymax = CumReturn, fill = "J203"), alpha = 0.3) +
    geom_ribbon(data = filter(FullIndex_df, Fund == "J403"), aes(ymin = 0, ymax = CumReturn, fill = "J403"), alpha = 0.3) +
    scale_color_manual(values = c("J203" = "blue", "J403" = "red")) +
    scale_fill_manual(values = c("J203" = "lightblue", "J403" = "pink")) +
    guides(fill = guide_none()) +
    scale_x_date(date_labels = "%Y-%m-%d") +
    labs(title = "Cumulative Returns by Index", x = "Date", y = "Cumulative Return") +
    theme_minimal()+
    facet_wrap(~ Index, scales = "free_y", ncol = 1)+
    theme(legend.position = "bottom")

finplot(index_plot)
```
When looking at Sectors, the SWIX(J403) and ALSI(J203) perform nearly identical when looking at the financial and property sector. ALSI outperforms SWIX in the industrial sector, but SWIX does better in resources. When looking at Index, interestingly, the ALSI outperforms SWIX across the board.

## Stratification

```{r warning=FALSE}
ZAR_returns <- ZAR %>% 
    filter(date > '2013-01-02') %>% #Filter for ZAR data after ALSI data start date
    mutate(Return = value/lag(value)-1) %>% 
    filter(date > first(date)) %>% 
    select(c(date, Return))


ZAR_sd <- ZAR_returns %>% 
    mutate(Year = format(date, '%Y')) %>% 
    group_by(Year) %>% 
    summarise(SD = sd(Return)*sqrt(12)) %>% 
  
    mutate(TopQuantile = quantile(SD, 0.8), BotQuantile = quantile(SD, 0.2))

Hi_Vol <- ZAR_sd %>%  filter(SD > TopQuantile) %>% pull(Year)
Low_Vol <- ZAR_sd %>%  filter(SD < BotQuantile) %>% pull(Year)


ReturnsDF <- calc_portfolio_returns(ALSI) %>% 
    mutate(Year= format(date, "%Y")) %>% 
    rename(Tickers = Fund, Return = Returns) %>% 
    group_by(Tickers) %>% 
    mutate(Top = quantile(Return, 0.99), Bot = quantile(Return, 0.01)) %>% 
    mutate(Return = ifelse(Return > Top, Top, 
                         ifelse(Return < Bot, Bot, Return))) %>% ungroup()

perf_hi <- Perf_comparisons(ReturnsDF, Ys = Hi_Vol, Alias = "High_Vol")
perf_lo <- Perf_comparisons(ReturnsDF, Ys = Low_Vol, Alias = "Low_Vol")

library(kableExtra)
RatioTable_High <- kable(perf_hi)
RatioTable_Low <- kable(perf_lo)
RatioTable_High
RatioTable_Low
```

Ratios above one indicate that that Ticker has a high (above usual) volatility for that given period. In the case of high volatility of the Rand, both the ALSI and SWIX show above usual volatility. Interestingly, ALSI also shows above usual volatility during periods of low Rand volatility.





```{r Figure2,  warning =  FALSE, fig.align = 'center', fig.cap = "Rolling 3 Year Annualized Returns SWIX vs ALSI (5% cap)", fig.ext = 'png', fig.height = 3, fig.width = 6}

RebDays <- RebDays %>% 
    filter(Date_Type == 'Reb Trade Day')

calculate_capped_port <- function(dataframe, W_cap, ColName) {

    Capped_df <- dataframe %>% 
        group_split(RebalanceTime) %>% 
        map_df(~Proportional_Cap_Foo(., W_Cap = W_cap)) %>% 
        select(-RebalanceTime)

    df_weights <- Capped_df %>% 
        tbl_xts(cols_to_xts = weight, spread_by = Tickers)

    df_returns <- dataframe %>% 
        filter(Tickers %in% unique(Capped_df$Tickers)) %>% 
        tbl_xts(cols_to_xts = Return, spread_by = Tickers)

    df_weights[is.na(df_weights)] <- 0
    df_returns[is.na(df_returns)] <- 0

    df_capped <- rmsfuns::Safe_Return.portfolio(R = df_returns, 
                                                  weights = df_weights, 
                                                  lag_weights = TRUE) %>% 
        xts_tbl()
    
    colnames(df_capped) <- c('date', ColName)

    return(df_capped)
}

### ALSI
rebalance_ALSI <- ALSI %>% 
    filter(date %in% RebDays$date) %>% 
    mutate(RebalanceTime = format(date, "%Y%B")) %>% 
    select(date, Tickers, Return, J203, RebalanceTime) %>% 
    rename(weight = J203) %>% 
    mutate(weight = coalesce(weight , 0))

ALSI_capped_10 <- calculate_capped_port(rebalance_ALSI, 0.1, 'ALSI_10')
ALSI_capped_5 <- calculate_capped_port(rebalance_ALSI, 0.05, "ALSI_5")

ALSI_10_rollrets <- ALSI_capped_10 %>% 
    mutate(RollRets_ALSI_10 = RcppRoll::roll_prod(1 + ALSI_10, 4, fill = NA, align = "right")^(4/12)-1) %>% 
    group_by(date) %>% 
    filter(any(!is.na(RollRets_ALSI_10))) %>% 
    ungroup() %>% 
    select(date, RollRets_ALSI_10)

ALSI_5_rollrets <- ALSI_capped_5 %>% 
    mutate(RollRets_ALSI_5 = RcppRoll::roll_prod(1 + ALSI_5, 4, fill = NA, align = "right")^(4/12)-1) %>% 
    group_by(date) %>% 
    filter(any(!is.na(RollRets_ALSI_5))) %>% 
    ungroup() %>% 
    select(date, RollRets_ALSI_5)



###SWIX
rebalance_SWIX <- ALSI %>% 
    filter(date %in% RebDays$date) %>% 
    mutate(RebalanceTime = format(date, "%Y%B")) %>% 
    select(date, Tickers, Return, J403, RebalanceTime) %>% 
    rename(weight = J403) %>% 
    mutate(weight = coalesce(weight , 0))
  
SWIX_capped_10 <- calculate_capped_port(rebalance_SWIX, 0.1, 'SWIX_10')
SWIX_capped_5 <- calculate_capped_port(rebalance_SWIX, 0.05, 'SWIX_5')

SWIX_10_rollrets <- SWIX_capped_10 %>% 
    mutate(RollRets_SWIX_10 = RcppRoll::roll_prod(1 + SWIX_10, 4, fill = NA, align = "right")^(4/12)-1) %>% 
    group_by(date) %>% 
    filter(any(!is.na(RollRets_SWIX_10))) %>% 
    ungroup() %>% 
    select(date, RollRets_SWIX_10)

SWIX_5_rollrets <- SWIX_capped_5 %>% 
    mutate(RollRets_SWIX_5 = RcppRoll::roll_prod(1 + SWIX_5, 4, fill = NA, align = "right")^(4/12)-1) %>% 
    group_by(date) %>% 
    filter(any(!is.na(RollRets_SWIX_5))) %>% 
    ungroup() %>% 
    select(date, RollRets_SWIX_5)


cap5_df <- left_join(ALSI_5_rollrets, SWIX_5_rollrets, by = 'date')
plotdf_5 <- cap5_df %>% 
    pivot_longer(-date,
                 names_to = 'Fund',
                 values_to = 'RollRet')

cap10_df <- left_join(ALSI_10_rollrets, SWIX_10_rollrets, by = 'date')
plotdf_10 <- cap10_df %>% 
    pivot_longer(-date,
                 names_to = 'Fund',
                 values_to = 'RollRet')

g <- plotdf_5 %>% 
    ggplot() + 
    geom_line(aes(date, RollRet, color = Fund), alpha = 0.7, size = 1.25) + 
    labs(title = "Rolling 3 Year Annualized Returns SWIX vs ALSI (5% cap)", 
        subtitle = "", x = "", y = "Rolling 3 year Returns (Ann.)", 
        caption = "") + 
    theme_fmx(title.size = ggpts(30), subtitle.size = ggpts(5), caption.size = ggpts(25), CustomCaption = F) + 
    fmx_cols()+
    scale_fill_brewer(palette = "Accent") 

finplot(g, x.date.dist = "1 year", x.date.type = "%Y", x.vert = T, 
    y.pct = T, y.pct_acc = 1)



```

```{r Figure3,  warning =  FALSE, fig.align = 'center', fig.cap = "Rolling 3 Year Annualized Returns SWIX vs ALSI (10% cap)", fig.ext = 'png', fig.height = 3, fig.width = 6}
g2 <- plotdf_10 %>% 
    ggplot() + 
    geom_line(aes(date, RollRet, color = Fund), alpha = 0.7, size = 1.25) + 
    labs(title = "Rolling 3 Year Annualized Returns SWIX vs ALSI (10% cap)", 
        subtitle = "", x = "", y = "Rolling 3 year Returns (Ann.)", 
        caption = "") + 
    theme_fmx(title.size = ggpts(30), subtitle.size = ggpts(5), caption.size = ggpts(25), CustomCaption = F) + 
    fmx_cols()+
    scale_fill_brewer(palette = "Accent") 

finplot(g2, x.date.dist = "1 year", x.date.type = "%Y", x.vert = T, 
    y.pct = T, y.pct_acc = 1)
```


The ALSI and the SWIX's rolling 3 year annualized returns closely follow each other. There is not much difference between the two for both a 5% and 10% cap. Interestingly, when looking at the period leading up to and straight after Covid-19, ALSI outperforms SWIX during the periods of high return, but under-performs during the downswing. This gap widens when looking at 10% capped portfolios.